---
title: "casestudy"
output:
  html_document: default
  pdf_document: default
---

```{r echo=TRUE}
beers<-read.csv("Beers.csv")
breweries<-read.csv("Breweries.csv")
library(tidyverse)

```
# Purpose
This studies relates two data sets together (breweries and the beer). From the brewery & beer merged data, we look at the bitterness (IBU) and alcohol by volume (ABV) in relation to their brewwery state. We then will look further into the relationship between IBU and ABV and see if there is a correlation.



# Question1 This code looks at the breweris data set and counts each row within the data set by the state column and returns a count of how often each state appears. Since each row is a unique brewery this will tell us how many breweries we have in each state in our data set.
This code looks at the breweris data set and counts each row within the data set by the state column and returns a count of how often each state appears. Since each row is a unique brewery this will tell us how many breweries we have in each state in our data set.


```{r echo=TRUE}
count(breweries,breweries$State)
```

# Question2 Merge beer data with breweries data by brewery id. Print first 6 observations and the last six observations to check the merged file.
First we need to merge the data. The R code below first looks at the columns of each data set. We want to map the beer data to the brewery data on the brewery ID. From the colnames code we can see that this columns do not have the same name.

```{r echo=TRUE}
colnames(beers)
colnames(breweries)
colnames(beers)[colnames(beers)=="Brewery_id"] <- "Brew_ID"
colnames(beers)
master_beer <- merge(breweries,beers,by="Brew_ID")
str(master_beer)
head(master_beer, n=6)

```

# Question3 Report the number of NA’s in each column.
The is.na function looks for values that are null, blank, N/A within the master_beer data set. The colsums function takes the sumation of those N/A’s by column showing us a clean summary of what data is not present.
```{r}
colSums(is.na(master_beer))
```


# Question4 Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
The first tapply function takes the master_beer data set, looks at the ABV by state and stores the median into med.abv because that is what we have told the tapply to do. In the second tapply notice the code is the same except we now tell it to calculate the IBU value and save it into med.IBU.

```{r}
med.abv<-tapply(master_beer$ABV, master_beer$State, median)
med.IBU<-tapply(master_beer$IBU, master_beer$State, median)
barplot(med.abv)

barplot(med.IBU)
```

# Question5 Which state has the maximum alcoholic beer? Which state has the most bitter beer?
The R code below uses the function which.max and then the data we are looking at. The which function ranks the data variables and the max pulls out the max value and displays the row. Now we know which brewery has the highest ABV and IBU.

```{r}
master_beer[which.max(master_beer$ABV),]
master_beer[which.max(master_beer$IBU),]

```


#Question6 Summary statistics for ABV (Alcohol by volume) variable.
```{r}
summary(master_beer$ABV)
```

# Question7 Is there a relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
Here is our scatter plot and a regression line which comes from “geom_smooth(method = lm)”. The scatter plot is defined by labeling our x and y axis from master_beer.
```{r}
ggplot(data = master_beer, mapping = aes(x = master_beer$ABV, y = master_beer$IBU))+
  geom_point() +
  geom_smooth(method = lm) 
cor.test(master_beer$ABV,master_beer$IBU)
```
#Conclusion
With a p value of 2.2e-16 we can say we should reject the null hypothesis that the intercept is 0 which is further proved by the confidence interval being .64 to .69 at 95%. Also the r-squared correlation = .6706 squared = .45 which is a fairly strong relationship. It’s safe to say that as I increase either ABV or IBU the other will increase as well.

It’s also worth noting that 1005 beers out of 2410 do not have an IBU value, but there is still 1405 observations with IBU which is a large sample still.
