---
title: "MSDS6306 Case Study 1"
author: "John Rodgers, Adam Scheerer, Sara Zaheri"
date: "October 13, 2018"
output:
  html_document:
    keep_md: yes
---

# Purpose
This study analyzes two data sets related to beer (breweries and the beer). The number of breweries for each state will be totaled and then We will look at the bitterness (IBU) and alcohol by volume (ABV) per state for all breweries in that state. We then will look further into the relationship between IBU and ABV and see if there is a correlation.

# Load Data

```{r load-data, echo=TRUE}
# Load Beers data
Beers <- read.csv("Beers.csv")
# Load Breweries data
Breweries <- read.csv("Breweries.csv")
# Adjust brewery ID column of Beers dataset to match brewery ID column name of Breweries dataset
colnames(Beers) [colnames(Beers)=="Brewery_id"] <- "Brew_ID"
```

# Question 1 - How Many breweries are in each state?
The following table displays the count of breweries for each state in the United States and the District of Columbia.
```{r q1, echo=TRUE}
#load data.table package
library(data.table)
#generate datatable with count of breweries for each state
breweries.by.state <- data.table(table(Breweries$State))
#assign variable names to datatable
colnames(breweries.by.state) <- c("State","Count")
#load knitr package
library(knitr)
kable(breweries.by.state)
```

# Question 2 - Merge the two data files together and show the first 6 and last 6 rows of the data.
For further analysis, the 2 files will be merged.  The following two tables display the first 6 rows and last 6 rows of the new combined dataset.
```{r q2, echo=TRUE}
#Merge 2 dataframes by column "Brew_ID"
beer.data.merged <- merge(Beers,Breweries, by ="Brew_ID")
#Clarify naming of "Name" columns
colnames(beer.data.merged)[colnames(beer.data.merged)=="Name.x"] <- "Beer Name"
colnames(beer.data.merged)[colnames(beer.data.merged)=="Name.y"] <- "Brewery Name"

#Display first 6 lines
kable(head(beer.data.merged,6))
#Display last 6 lines
kable(tail(beer.data.merged,6))
```

# Question 3 - Display the number of NA values for each column
In order to identify missing values in the dataset, the following table displays the number of records with missing values for each of the following columns in the data.
```{r q3, echo=TRUE}
kable(colSums(is.na(beer.data.merged)), col.names=c("Count"))
```

# Question 4 - Compute and plot the median ABV and IBU for each state
In order to visually analyze the median ABV as compared to the median IBU for each state to determine if there is a relationship between the two variables, the following chart displays a barchart of the IBU values compared to a line chart of ABV values.  The scale for the ABV calculations is 1:10 of the scale for IBU, with the scale values displayed on the top for ABV and on the bottle for IBU
```{r q4, fig.height=10, echo=TRUE}
#Calculate median ABV by State
med.ABV<-aggregate(IBU ~ State, beer.data.merged, median)
#Calculate median IBU by State
med.IBU<-aggregate(ABV ~ State, beer.data.merged, median)
#Load reshape2
library(reshape2)
#Melt data for ggplot2
IBU <- melt(med.ABV)
ABV <- melt(med.IBU)
#Load ggplot2
library(ggplot2)
#Plot IBU and ABV
ggplot() +
  geom_bar(data=IBU, aes(x=State, y=value, fill=variable),stat = "identity") + 
  geom_line(data=ABV, aes(x=State, y=value*1000, colour=variable, group=1), stat="identity") +
  coord_flip() +
  theme(legend.key = element_blank(), 
        legend.background = element_rect(colour = 'black', fill = 'white'), 
        legend.position = "right", legend.title = element_blank(),
        legend.box.just = "left") +
  scale_y_continuous(sec.axis = sec_axis(~./10, name="Median ABV%")) +
  ylab("Median IBU") 
```

# Question 5 - Which state has the maximum ABV and which has the maximum IBU?
The beers with the highest ABV values and highest IBU values are displayed in the following tables.
```{r q5, echo=TRUE}
#Highest ABV
kable(beer.data.merged[which.max(beer.data.merged$ABV),c(2,4:10)], caption="Highest ABV")
#Highest IBU
kable(beer.data.merged[which.max(beer.data.merged$IBU),c(2,4:10)], caption="Highest IBU")
```

# Question 6 - Display the summary statistics of ABV
The following table displays a numerical analysis of of values of the ABV column.
``` {r q6, echo=TRUE}
summary(beer.data.merged$ABV)
```

#7 Question 7 - Display a scatterplot to show the relationship between bitterness and ABV
The scatterplot below displays further analysis of the relationship between IBU and BV.  The line and the shading around it represents a linear model and the confidence interval.  
```{r q7, echo=TRUE}
#Display scatterplot of ABV vs IBU, including linear model and confidence interval
ggplot(beer.data.merged, aes(x=IBU, y=ABV)) + geom_point() + geom_smooth(method=lm)
cor.test(beer.data.merged$ABV,beer.data.merged$IBU)
```

# Conclusion

With a p-value < .0001 there is significant evidence that the null hypothesis that there is a correlation between the ABV and IBU of beers. We can be 95% confident that the correlation between ABU and IBU is between .64 and .70. It is worth noting that 1005 beers out of 2410 do not have an IBU value, excluding them from this correlation analysis.